---
title: "Structural Equation Models with Latent Variables"
author: "Jeremy Leipzig"
date: "May 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data
```{r dataset, echo=FALSE}
debateMen<-'
21.1
18.5  21.4
19.3  19.2  21.1
 5.3   5.6   5.7   7.9
 5.0   4.8   5.2   5.2   6.5
 5.1   5.3   5.2   0.2   0.0   6.3
 5.4   5.6   5.3   0.1   0.0   4.7   6.7'

debateWomen<-'
39.4
37.1  39.2
37.9  37.6  40.5
11.1  11.2  11.4   7.8
10.0   9.8  10.3   5.2   6.9
 5.3   5.7   5.5   0.2   0.9   6.8
 5.6   6.1   5.5   0.1   0.0   4.5   6.7'

library(lavaan)
library(semPlot)
library(gridExtra)
debateMen.cov <- 
    getCov(debateMen, names = c("S1", "S2", 
                            "S3", "P1",
                            "P2", "Q1", "Q2"))
debateWomen.cov <- 
    getCov(debateWomen, names = c("S1", "S2", 
                            "S3", "P1",
                            "P2", "Q1", "Q2"))
```

```{r}
knitr::kable(debateMen.cov)
```

# Fair Model
I assumed the ratings would be a product of the personableness of the candidate and the quality of the arguments.

```{r formulation}



fair.model <- '
  # latent variables
    scales     =~ S1 + S2 + S3
    personable =~ P1 + P2
    argument =~ Q1 + Q2
  # regressions
    scales ~ personable + argument
'



  # regressions
    #alien71 ~ alien67 + ses
    #alien67 ~ ses
  # correlated residuals
    #anomia67 ~~ anomia71
    #powerless67 ~~ powerless71
fit.fair.men <- sem(fair.model, 
           sample.cov = debateMen.cov, 
           sample.nobs = 154)
fit.fair.women <- sem(fair.model, 
           sample.cov = debateWomen.cov, 
           sample.nobs = 125)
heatmap(debateMen.cov)
heatmap(debateWomen.cov)
```

```{r plotsem, echo=FALSE}
semPaths(fit.fair.men, title = FALSE, curvePivot = TRUE)
```

### Summary of males
```{r}
summary(fit.fair.men, standardized = TRUE, fit.measures = TRUE)
```

### Summary of females
```{r}
summary(fit.fair.women, standardized = TRUE, fit.measures = TRUE)
```

# Partisan Model
Here I predicted arguments would fall in line with personableness
```{r}
partisan.model <- '
  scales     =~ S1 + S2 + S3
    personable =~ P1 + P2
    argument =~ Q1 + Q2
    argument ~ personable
    scales ~ argument
'

fit.partisan.men <- sem(partisan.model, 
           sample.cov = debateMen.cov, 
           sample.nobs = 154)
fit.partisan.women <- sem(partisan.model, 
           sample.cov = debateWomen.cov, 
           sample.nobs = 125)
```

```{r partisanplotsem, echo=FALSE}
semPaths(fit.partisan.men, title = FALSE, curvePivot = TRUE)
```

### Summary of partisan males
```{r}
summary(fit.partisan.men, standardized = TRUE, fit.measures = TRUE)
```

### Summary of partisan females
```{r}
summary(fit.partisan.women, standardized = TRUE, fit.measures = TRUE)
```

### Comparison
```{r}
anova(fit.fair.men, fit.partisan.men)
anova(fit.fair.women, fit.partisan.women)
```

It appears the fair model is a better fit, and that personableness is a modestly stronger indicator than predictor of ratings for women than men.
